---
title: "Introduction to R: Problems"
output: html_notebook
---

# Numbers, vectors and basic functions
2.1 Create a vector "prob" with 50 elements which has the following form $prob = \begin{pmatrix} 1 & 4 & 9 & 16 & \cdots\end{pmatrix}$. 

```{r}
prob = seq(from=1, to=50, by=1)
prob = prob^2
print(prob)
```


2.2 Calculate the scalarproduct for the vector prob with itself.

```{r}
prob_scalar = prob%*%prob
print(prob_scalar)
```


2.3 Take the square root of the vector prob and add this to this vector. How can you obtain the same result? Test whether the difference is indeed 0.

```{r}
prob_sqrt = sqrt(prob)
prob_transform1 = prob + prob_sqrt
prob_transform2 = prob + seq(from=1, to=50, by=1)
identical(prob_transform1, prob_transform2)
```


2.4 Construct another vector "alphabet" consisting of characters. To this end use the build in variable letters. Can you add the vectors alphabet and prob? Why or why not? Set up a logical vector "logi"  and use it instead of alphabet. What can you conclude?

```{r}
alphabet = letters
print(alphabet)

# Test prob + alphabet
# prob + alphabet # Afterwards, delete this line or turn it into a comment to continue
logi = rep(c(TRUE, FALSE), 25)
prob + logi
```
You cannot add a vector of whose elements are character type ones to a numeric vector. Additionally, the dimensions would not match.
Assuming that logi has the same length as prob, TRUE is interpreted as 1 while FALSE is treated as 0.


2.5 Find the minimal and maximal value as well as the sum and the range of the vector zeta. In addition use the function \emph{summary()}.

```{r}
zeta = rnorm(1000)
zeta_min = min(zeta)
zeta_max = max(zeta)
zeta_sum = sum(zeta)
zeta_mean = mean(zeta)
zeta_range = range(zeta)
print(paste("zeta_min: ", zeta_min))
print(paste("zeta_max: ", zeta_max))
print(paste("zeta_sum: ", zeta_sum))
print(paste("zeta_mean: ", zeta_mean))
print(paste("zeta_range: min =", zeta_range[1], ", max =", zeta_range[2], ", difference = ", zeta_range[2]-zeta_range[1]))

summary(zeta)


```


2.6 Extract now the minimal and maximal values of the vector zeta and save these in the variables min and max. Additionally count how many times you obtained a value smaller than 0. How many datapoints are in the interval from -1 to 1?

```{r}
min = zeta_min
max = zeta_max
smaller_zero = sum(zeta < 0)
minus_one_to_one = sum(zeta >= -1 & zeta <= 1)
print(paste("No. of elements < 0: ", smaller_zero))
print(paste("No. of elements -1 < zeta < 1: ", minus_one_to_one))
```


2.7 Construct a numeric variable "threshold" with any number you like. Exclude all elements of the vector zeta that are greater than your threshold value. Repeat this procedure with different threshold values.

```{r}
zeta_sort = sort(zeta)
zeta_threshold = zeta[zeta >=  -1 & zeta <= 1]
range(zeta_threshold)
length(zeta_threshold)
```


# Logical statements, loops and matrices.
4.1 Use a for-loop to calculate the mean of the vector zeta. Test whether your result is correct. Repeat this using a while-loop.

```{r}
# Create some vector zeta of length n
n = 1000
# zeta = rnorm(n)

# Calculate the mean of zeta using a for loop
mean_for = 0
for (i in 1:n){
  mean_for = zeta[i]/n + mean_for
}
# Test the calculation
identical(mean_for, mean(zeta))

# Calculate the mean of zeta using a while loop
i = 1
mean_while = 0
while (i <= n){
  mean_while = zeta[i]/n + mean_while
  i = i+1
}
identical(mean_while, mean(zeta))
```


4.2 Write a function to calculate the double factorial $n!! = \prod\limits_{k=0}^{\lceil \frac{n}{2} \rceil - 1}(n-2k)$ of an integer using a for- or while-loop. Check within the function whether the integer is not negative, if not return an error message. Hint: The function _ceiling()_ might be useful.

```{r}
doubleFac = function(n){
  # Inputs:
  #   n: Positive integer
  # Outputs:
  #   res: res = n!!
  
  if (n > 0){
    prod = 1
    for (k in seq(0, ceiling(n/2)-1)){
      prod = prod*(n-2*k)
    }
    return(prod)
  }
  else if (n == 0){
    return(1)
  }
  else if (n < 0){
    return("n must be positive!")
  }
}
doubleFac(-1)
doubleFac(0)
doubleFac(9)
```


4.3 Write a function to calculate the greatest common divisor for two positive integers using a while loop. To this end use the euclidean algorithm. Check within the function whether the integers are positive, if not return an error message. Hint: y modulus x is _y%%x_ in R.

```{r}
gcd = function(a, b){
  # Inputs
  #   a, b: Positive numbers
  # Outputs
  #   d: Greatest common divisor of a and b
  
  # Euclid's Algorithm
  d = 1
  if ((a>0) & (b>0)){
    while(b != 0){
      t = b
      b = a%%b
      a = t
    }
    d = a
    return(d)
  }
  else{
    return("At least one number is not positive.")
  }
}

gcd(5,2)
gcd(2,4)
gcd(-2,5)
```


4.4 Use the vectors a1 and b1 to verify both DeMorgan's laws. Do not forget to test the result!

```{r}
a1 = rbinom(n = 10000, size = 1, prob = 0.5)
b1 = rbinom(n = 10000, size = 1, prob = 0.5)
all(!(a1|b1)==((!a1)&(!b1))) 
all(!(a1&b1)==((!a1)|(!b1)))
```


4.5 Write a procedure to obtain the smallest value in the vector zeta. Use a for- or while-loop. Compare your procedure to the function _min()_.

```{r}
# Test vector
zeta = rnorm(1000)

# Create a function
min_vec = function(x){
  # Inputs
  #   x: some vector
  # Outputs
  #   res: Minimum of vector x
  
  n = length(x)
  res = x[1]
  for (i in 2:n){
    if (x[i] < res){
      res = x[i]
    }
  }
  return(res)
}
# Test the function and compare the result to min(zeta)
min_vec(zeta)
identical(min_vec(zeta), min(zeta))
```

4.6 Write a function that creates an $n\times n$ identity matrix using for- or while-loops. n is the only argument of this function.

```{r}
identity = function(n){
  # Inputs:
  #   n: Dimension of the square identity matrix
  # Outputs:
  #   nxn identity matrix
  
  iden = matrix(0, nrow=n, ncol=n)
  for (i in 1:n){
    iden[i, i]=1
  }
  return(iden)
}
identity(2)
identity(10)
```

4.7 Write a function that creates an $n\times n$ matrix of the following form $\begin{pmatrix} 1 & 2 & 3 & \cdots & n\\ 2 & 2 & 3 & \cdots & n\\ 3 & 3 & 3 & \cdots & n\\ \vdots & \vdots & \vdots & \ddots & \vdots\\ n & n & n & \cdots & n\end{pmatrix}$. n is the only argument of this function.

```{r}
mat = function(n){
  # Inputs:
  #   n: Dimension of the matrix
  # Output
  #   mat_res: nxn matrix
  
  mat_res = matrix(0, nrow=n, ncol=n)
  for (i in 1:n){
    for (j in 1:n){
      if (i >= j){
        mat_res[i, j] = i
      }
      else{
        mat_res[i, j] = j
      }
    }
  }
  return(mat_res)
}

# mat()
mat(5)
mat(10)
```