---
title: 'Introduction to R: Problems'
output: html_notebook
---

# Numbers, vectors and basic functions
2.1 Create a vector "prob" with 50 elements which has the following form $prob = \begin{pmatrix} 1 & 4 & 9 & 16 & \cdots\end{pmatrix}$. 

```{r}

```


2.2 Calculate the scalarproduct for the vector prob with itself.

```{r}

```


2.3 Take the square root of the vector prob and add this to this vector. How can you obtain the same result? Test whether the difference is indeed 0.

```{r}

```


2.4 Construct another vector "alphabet" consisting of characters. To this end use the build in variable letters. Can you add the vectors alphabet and prob? Why or why not? Set up a logical vector "logi"  and use it instead of alphabet. What can you conclude?

```{r}

```


2.5 Find the minimal and maximal value as well as the sum and the range of the vector zeta. In addition use the function \emph{summary()}.

```{r}
zeta = rnorm(1000)
```


2.6 Extract now the minimal and maximal values of the vector zeta and save these in the variables min and max. Additionally count how many times you obtained a value smaller than 0. How many datapoints are in the interval from -1 to 1?

```{r}

```


2.7 Construct a numeric variable "threshold" with any number you like. Exclude all elements of the vector zeta that are greater than your threshold value. Repeat this procedure with different threshold values.

```{r}

```


# Logical statements, loops and matrices.
4.1 Use a for-loop to calculate the mean of the vector zeta. Test whether your result is correct. Repeat this using a while-loop.

```{r}

```


4.2 Write a function to calculate the double factorial of an integer using a for- or while-loop. Check within the function whether the integer is not negative, if not return an error message. Hint: The function _ceiling()_ might be useful.

```{r}

```


4.3 Write a function to calculate the greatest common divisor for two positive integers using a while loop. To this end use the euclidean algorithm. Check within the function whether the integers are positive, if not return an error message. Hint: y modulus x is _y%%x_ in R.

```{r}

```


4.4 Use the vectors a1 and b1 to verify both DeMorgan's laws. Do not forget to test the result!

```{r}
a1 = rbinom(n = 10000, size = 1, prob = 0.5)
b1 = rbinom(n = 10000, size = 1, prob = 0.5)
```


4.5 Write a procedure to obtain the smallest value in the vector zeta. Use a for- or while-loop. Compare your procedure to the function _min()_.

```{r}
zeta = rnorm(1000)
```

4.6 Write a function that creates an $n\times n$ identity matrix using for- or while-loops. n is the only argument of this function.

```{r}

```

4.7 Write a function that creates an $n\times n$ matrix of the following form $\begin{pmatrix} 1 & 2 & 3 & \cdots & n\\ 2 & 2 & 3 & \cdots & n\\ 3 & 3 & 3 & \cdots & n\\ \vdots & \vdots & \vdots & \ddots & \vdots\\ n & n & n & \cdots & n\end{pmatrix}$. n is the only argument of this function.