---
title: "Analyzing data"
output: html_notebook
---

## Description of the variables
The dataset consists of the following variables:

Name       | Description
-----------|------------
net_inc    | monthly net income
life_sat   | satisfaction with life on a scale from 0 (very bad) to 10 (very good)
state      | state of residence in Germany
doc_visits | number of doctor visits in a year
bf_extra   | Extraversion
bf_agree   | Agreeableness
bf_neuro   | Neuroticism
bf_rigid   | Conscientiousness
bf_open    | Openess to experience
educ       | education 
age        | age of participant

The big five personality traits are measured on a scale from 0 to 5 each.

For education we have the following possible outcomes:

value     | Description
----------|------------
0         | no graduation
1         | lower or intermediate secondary school
2         | high school
3         | college, university or specialized college of higher education

## Loading data
Load the data in example.dta in R. The library *foreign* might be helpful.

```{r}
library(foreign)
data = read.dta("example.dta") #Make sure your working directory is set

```

## Descriptive statistics and basic plots
Calculate descriptive statistics of all variables ( *summary()* ), as well as a frequency chart for satisfaction (*barplot()*), and box plots (*boxplot()*) for both number of doctor visits and net income. Additionally construct a single figure containing both box plots and save this in a pdf file (*pdf()*).

```{r}
# Get the descriptive statistics for the dataset
summary(data)


# Plot the bar chart for satisfaction
summary(data$life_sat)
barplot(table(data$life_sat), names.arg = seq(-6, 10, 1), xlab="Life_sat factors", ylab="# of observations", main="Observed life satisfaction")

# Plot the boxplots
par(mfrow = c(1, 2)) # Subplot setting
boxplot(data$doc_visits, ylab="# Doctor visits")
boxplot(data$net_inc, ylab="Monthly net income")
par(mfrow = c(1, 1)) # Reset subplot setting

# Create and export a pdf
file_path = file.path(getwd(),"Boxplots.pdf")
pdf(file = file_path)
par(mfrow = c(1,2))
boxplot(data$doc_visits, ylab = "number of doctor visits")
boxplot(data$net_inc, ylab = "monthly net income")
dev.off() # Stop storing to the pdf
par(mfrow = c(1,1)) # reset the plot grid
```


## Transformations
Add age^2 to the current dataset. Remove all missing values and for the categorical variables satisfaction and state of residence delete additionally all observations where participants haven't answered (*na.omit()*). Change education to a factor variable (*factor()*).

```{r}
# Add age^2
data$age2 = data$age^2

# Remove missing arguments
data = na.omit(data)
# Save which observations were removed from the data
data_missing_remove = na.action(na.omit(data))

# Remove those who did not answer life_sat or state
data$life_sat = factor(data$life_sat, exclude = c(NA, "[-1] No Answer"))#, labels = 0:10)
data$state = factor(data$state, exclude = c(NA, "[-1] No Answer"))

# Change education to a factor variable
data$educ = factor(data$educ)

# Remove obviously wrong observations (wrongly answered - maybe on purpose, maybe wrongly stored)
for (i in 1:dim(data)[1]){
  if (data$net_inc[i]<0){
    data$net_inc[i] = NA
  }
  if (data$doc_visits[i]<0){
    data$doc_visits[i] = NA
  }
} # Replace 

data = na.omit(data)
data_answer_remove = na.action(na.omit(data))
```

## Estimation
We are interested in the determinants of (net-) income, therefore specify a linear model and estimate it by simple OLS (*lm()*). Save the regression results in a file.

```{r}
# Build the regression formula
# Grab the endogenous variable
response = colnames(data)[1]
# Grab the exogenous variables
variables = colnames(data)[2:length(colnames(data))]
# Put them together as a string of formula type
formula = as.formula(
  paste(response,
        paste(variables, collapse = " + "),
  sep = " ~ ")
)

# Fit the simple linear model
fit = lm(formula=formula, data=data)
summary(fit)

# Store the results in a txt document
out = capture.output(summary(fit))
cat("Estimation results:", out, file = "Results.txt", sep = "\n") # append = T
```

## Prediction
Predict (*predict()*)the net income of a 25 year old individual with a bachelor degree living in SH which visits the doctor 3 times a year. All big five traits as well as satisfaction can be assumed to be average.

```{r}
x1 = "[5] 5 Satisfied: On Scale 0-Low to 10-High"
x2 = "[1] Schleswig-Holstein       1"
x3 = 3
x4 = 3
x5 = 3
x6 = 3
x7 = 3
x8 = 3
x9 = "3"
x10 = 25
x11 = 25^2
forecast = data.frame(x1, x2, x3, x4, x5, x6, x7, x8, x9, x10, x11)
colnames(forecast) = variables
predict(object = fit, newdata = forecast)
```

