---
title: "Tutorial 5"
output: html_document
---

## Convergence in distribution
### a)
Binomial distribution for large n compared to a Poisson distribution. 

```{r}

```

### b) 
Student-t distribution with many degrees of freedom compared to Standard Normal distribution.

```{r}

```

### c) 
Binomial distribution B(n,p) with large n compared to a Normal distribution with mean np and variance np(1-p), (np(1-p) > 9 or n >= 100).

```{r}

```

Plot the result using a histogram with Scott breaks. 

```{r}

```

## Stutsky's theorems:
Let $X_n \overset{d}{\rightarrow} X$ and $Y_n \overset{p}{\rightarrow} c$, then
1. $X_n + Y_n \overset{d}{\rightarrow} X + c$.
2. $X_n \cdot Y_n \overset{d}{\rightarrow} X \cdot c$.
3. $\frac{X_n}{Y_n} \overset{d}{\rightarrow} \frac{X}{c}$ if $P(Y_n=0)=0$ and $c \neq 0$.

Assume $X_n \sim \mathcal{N} \left(\mu + \frac{1}{n}, \sigma^2\right)$ and $Y_n \sim \mathcal{N} \left(\mu, \frac{\sigma^2}{n}\right)$. 
Verify Slutsky's theorems with the help of Montee Carlo simulations. Choose $N_{MC}=1000$, $n=1000$.

```{r} 

```

```{r}

```

```{r}

```


