---
title: "Analyzing data"
output: html_notebook
---

## Description of the variables
The dataset consists of the following variables:

Name       | Description
-----------|------------
net_inc    | monthly net income
life_sat   | satisfaction with life on a scale from 0 (very bad) to 10 (very good)
state      | state of residence in Germany
doc_visits | number of doctor visits in a year
bf_extra   | Extraversion
bf_agree   | Agreeableness
bf_neuro   | Neuroticism
bf_rigid   | Conscientiousness
bf_open    | Openness to experience
educ       | education 
age        | age of participant

The big five personality traits are measured on a scale from 0 to 5 each.

For education we have the following possible outcomes:

value     | Description
----------|------------
0         | no graduation
1         | lower or intermediate secondary school
2         | high school
3         | college, university or specialized college of higher education

## Loading data
Load the data in example.dta in R. The library *foreign* might be helpful.

```{r}

```

## Descriptive statistics and basic plots
Calculate descriptive statistics of all variables ( *summary()* ), as well as a frequency chart for satisfaction ( *barplot()* ), and box plots ( *boxplot()* ) for both number of doctor visits and net income. Additionally construct a single figure containing both box plots and save this in a pdf file ( *pdf()* ).

```{r}

```


## Transformations
Add age^2 to the current dataset. Remove all missing values and for the categorical variables satisfaction and state of residence delete additionally all observations where participants haven't answered ( *na.omit()* ). Change education to a factor variable ( *factor()* ).

```{r}

```

## Estimation
We are interested in the determinants of (net-) income, therefore specify a linear model and estimate it by simple OLS ( *lm()* ). Save the regression results in a file.

```{r}

```

## Prediction
Predict ( *predict()* ) the net income of a 25 year old individual with a bachelor degree living in SH which visits the doctor 3 times a year. All big five traits as well as satisfaction can be assumed to be average.

```{r}

```

