---
title: "HA_Prob_Calc_Winter23_24"
author: "Rouven Lindenau"
date: "2024-01-22"
output: html_document
---
---
title: "Home Assignment, Probability Calculus 2023/2024"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Procedure
- Try to solve this home assignment and answer all questions in this file by filling the boxes below each exercise
```{r}
# Example Code here
print("It works!")
```
- You are not allowed to use any additional R packages (this means packages that you have to install additionally to those already installed by default). 
- You can work in groups of up to 5 students. 
- Hand in your solution (as a Rmd file!) via [e-mail] (l.baumann@stat-econ.uni-kiel.de) and (rl@stat-econ.uni-kiel.de) until Sunday, 11.02.2024, 23:59:59. 
- Make sure to include all names and stu-numbers in your mail and in the header of this file!
- We won't evaluate your assignment if your code is not running!
- Add some comments to your code lines so people can understand your code more easily.
- We might ask you to explain your code if we suspect that you did not write it on your own.


!!! Enter your group name and stu-numbers here !!!

1 Marque Mollenhauer, stu227420
2 Eric Kroll, stu243616
3 Josef Fella, stu245231
4 Robert Hennings, stu236320
5 Ahsan Muhammad, stu243716

!!! Enter your group name and stu-numbers here !!!

_______________________________________________________________________________________________________________
#1. a) Exercise: Password generator (3 Points)

You are asked to write a function that creates random passwords. The function must have two arguments:
1. length - number of symbols used in the password, it should be 4 by default if no value is specified.
2. special - a TRUE/FALSE boolean that is TRUE if special characters like "$" are included, it should be TRUE by default, if no value is specified.

The password can be made of 
- lower case letters, "a", "b",...,"z" (there are 26 possibilities)
- upper case letters, "A", "B",...,"Z" (there are 26 possibilities)
- numbers, "0", "1",...,"9" (there are 10 possibilities)
- and (if allowed) special characters which are: "$", "%", "&", "/", "(", ")" (there are 6 possibilities)

Further, the password is supposed to have at least one of each: lower case letter, upper case letter, number and if allowed special character. 
Apart from that, at each position the password should have an equal probability to draw one of the 68 symbols (62 if special characters not allowed).

Additionally, 
3. If "length" in the function argument is less than 4, return: "Password must be at least of length 4!"
4. If "length" in the function argument is not an integer, return: "Number of symbols must be a positive integer!"

Last,
5. If length of the password is <6, the function should print: "Weak password!"
6. If length of the password is >6 and <10, the function should print: "Moderate password!"
7. If length of the password is >10, the function should print: "Strong password!"

8. The function should return the password as one single character, i.e. merging the drawn symbols in one character.

Make sure your function fulfills all needed conditions by testing a few variations!

```{r}
# Write your function here, Conditions 1) and 2)
get_password <- function(length=4, special=TRUE){
  options(warn=-1)
  # Test for type of provided argument - taken from the R help directly
  # see help page for integer, bottom part
  is.wholenumber <-
    function(x, tol = .Machine$double.eps^0.5)  abs(x - round(x)) < tol
  # Test for length of length var - Condition 3)
  if (length < 4) {
    return("Password must be at least of length 4!")
  }
  # Test for datatype of argument - Condition 4)
  if (is.wholenumber(length)==F){
    return("Number of symbols must be a positive integer!")
  }
  # at least one lower, upper, number, if allowed special work with modulo operator
  # and distribute the rest across the symbol categories to draw from
  specials <- c("$", "%", "&", "/", "(", ")")
  numbers <- 0:9
  all_symbols <- c(LETTERS, letters, numbers)
  # Always draw the standard symbols
  first_pw <- c(sample(LETTERS, size=1),
                sample(letters, size=1),
                sample(numbers, size=1))
  
  if (special == TRUE) {
    # include draws directly from the special characters to fill up the rest
    first_pw <- c(first_pw, sample(specials, size = 1))
    
    pw <- c(first_pw, sample(all_symbols, size=length-length(first_pw)))
    
  } else {
    # if there's rest, it is now drawn only from all standard symbols
    pw <- c(first_pw, sample(all_symbols, size=length-length(first_pw)))
    }
  # Password Strength - nested ifelse - Conditions 5), 6), 7)
  # Not really sure here, maybe typo in exercise? shouldn't it be >=6 &<10?
  ifelse(length<6, print("Weak password!"),
         ifelse(10>=length & length>=6, print("Moderate password!"),
                print("Strong password!")))
  # A final sample to mix up the order and extract spaces - Condition 8)
  pw <- paste0(sample(pw, size=length(pw)), collapse="")
  return(pw)
}
# Test function
get_password(length=4, special=F)
```

#1. b) Test your function here for

```{r}

length = 7
special = T
# your function executed here
get_password(length = length, special = special)

length = 3
special = T
# your function executed here
get_password(length = length, special = special)

length = 4.5
special = T
# your function executed here
get_password(length = length, special = special)

length = 5
special = F
# your function executed here
get_password(length = length, special = special)

length = 50
special = T
# your function executed here
get_password(length = length, special = special)
```


# 2. a) Excercise: Gamma vs. Normal distribution (3 points):

In this exercise you need to demonstrate that if a random variable $X$ is ${\Gamma}$-distributed with shape parameter $a$ and scale parameter $b$, then for $a \to \infty$  $X \overset{d}{\rightarrow} \mathcal{N} (ab,ab^2)$ if $a > b$.

To this end, write a function that samples and returns $N$ observations from the ${\Gamma}$ distribution with shape parameter $a$ and scale parameter $b$. The function's input shall be $a$, $b$ and N.

```{r}
# Inspect Gamma samples and plot vs. the limiting normal distribution
N <- 5000
a <- 1
b <- 1
breaks <- "Scott"
# Create the Plot and compare the gamma samples with the limiting normal distribution
get_gamma_vs_normal <- function(N, a, b, breaks){
  gam_samp <- rgamma(n=N, shape=a, rate=b)  
  x_ <- seq(min(gam_samp), max(gam_samp), length.out=length(gam_samp))
  hist(gam_samp, probability=T, breaks=breaks,
       main=sprintf("N: %s simulated random Gamma values for a: %s, b: %s", N, a, b),
       xlab="x-values")
  lines(x_, dnorm(x_, mean=a*b, sd=sqrt(a*b^2)), col="red")
  legend("topright", legend=c("Norm PDF"), col=c("red"), lty=c(1), lwd = c(1))
}
get_gamma_vs_normal(N, a, b, breaks)
```

Plot the resulting observations from the ${\Gamma}$ distribution using histograms with Scott breaks for $a$ = c(1, 5, 10, 50, 100) and $b$ = 1, N = 5000 in a 3x2 plot window and add the respective limiting Normal density for comparison. Choose a meaningful title, label the axes and add a legend to your plot.

```{r}
a <- c(1, 5, 10, 50, 100, 10000)
b <- 1
breaks <- "Scott"
N <- 5000

get_gamma_normal_plot <- function(a, b, breaks, N) {
  options(warn=-1)
  # 3x2 plot
  nf <- layout(matrix(c(1,2,3,4,5,6), ncol=2,nrow=3, TRUE),widths = c(1,1), heights=c(1,1,1))
  par(mar = c(2,4,1,0) + 0.1)
  for (a_ in a) {
    gamma_sample <- rgamma(N, shape = a_, scale = b)
    x <- seq(min(gamma_sample), max(gamma_sample), length.out = N)
    # Plot Gamma values against limiting normal distribution given as Hint from the Task
    norm_dist_pdf <- dnorm(x, mean = a_ * b, sd = sqrt(a_ * b^2))
    
    hist(gamma_sample, breaks = breaks, probability = TRUE,
         main = sprintf("Gamma Samples (a:%s, b:%s) vs. Lim Normal Distribution", a_, b),
         xlab = "x values", ylab = "Density", cex.main = 1)
    lines(x = x, y = norm_dist_pdf, type = "l", col = "red")
    # Additionally plot the mean of the Gamma and the normal distribution
    abline(v=mean(gamma_sample), col="black", lwd=1)
    abline(v=a_ * b, col="blue", lwd=1)
    legend("topright",
           legend = c("Norm PDF", "Mean Gamma", "Mean Norm"),
           col = c("red", "black", "blue"),
           lty = c(1, 1, 1),
           lwd=c(1, 1, 1))
  }
  par(mfrow = c(1, 1))
}

get_gamma_normal_plot(a=a, b=b, breaks=breaks, N=N)
```

# 2. b) Slutsky's theorems

Now assume that you observe $X_n \sim \Gamma(a,b)$ for $a = c(1, 5, 10, 50, 100)$ and $b = 1$ as well as $Y_n \sim \mathcal{N}\left(\mu, 4\sigma^2\sqrt{n^{-5}}\right)$ where $\mu = 3.5$ and $\sigma = 7$. Verify that following part of Slutsky's theorems holds with increasing values of $a$ : 2. $X_n \cdot Y_n \overset{d}{\rightarrow} X\cdot c$. Do this with the help of Monte Carlo simulations and a corresponding 3x2 plot window. Choose $N_{MC} = 1000$, $n = 1000$.
Hint: Remember that $X \overset{d}{\rightarrow} \mathcal{N} (ab,ab^2)$ for increasing values of $a$.

```{r}
# Settings
N_MC <- 1000
n <- 1000
a <- c(1, 5, 10, 50, 100, 10000)
b <- 1
mean_y <- 3.5
sigma_y <- 7
sd_y <- sqrt((4*sigma_y^2)*sqrt(n^-5))
c <- mean_y
# Theoretical Ideas:
# The Normal distribution is related to the Gamma Distribution, so if we multiply something related to a normal distribution with something truly normal the result should be as well something normally distributed
# By the Hint provided, that the X follows: $X \overset{d}{\rightarrow} \mathcal{N} (ab,ab^2)$ for increasing a, we can calculate at each a the needed parameters for the given distribution as it can be seen as a function of a and, respectively a constant b set to 1, but we must not forget to scale the limiting distribution accordingly by the constant c, which we can not choose arbitrarily. Since the X part is more or less the same on the left and right hand side, the only difference is the constant factor c on the right vs the distribution of Y on the left hand side, so this distribution has to be equal to the factor c in its limit, we assume that its converging in the long run towards its expected value what is the given value for mu which is 3.5, so c has to be equal to mu which is 3.5, this can be seen letting lim n->inf where the sd of Y converges towards 0 leaving only the of n independent mu as resulting value

get_comp_plot <- function(a, b, c, n, N_MC, mean_y, sd_y){
  # First simulate X and Y and combine them to one single factor
  get_x_y_n <- function(a, b, n, N_MC){
    # simulate N_MC * n values, N_MC runs of the size of n
    gamma_samp <- matrix(rgamma(n=n * N_MC, shape=a, scale=b), ncol = N_MC)
    normal_samp <- matrix(rnorm(n=n * N_MC, mean=mean_y, sd=sd_y), ncol = N_MC)
    x_y_n <- gamma_samp * normal_samp
    return(x_y_n)
  }
  # Next get the factor X*c
  get_x_c <- function(a, b, c){
    # From the given Hint we know that X follows a Normal(ab, ab^2) distribution
    # which we have to change a little bit to also incorporate the constant factor c
    # Here the number of simulations isnt playing a role since we only need the few fixed
    # values for coputation
    X_mean <- a*b
    X_var <- a*b^2
    # Here also slightly change the Hint by multiplying the mean with the constant c
    # Orientation provides Programming Tutorial 5) Slutzkys Theorem
    X_mu <- X_mean * c
    # Here also slightly change the Hint by multiplying the sd with the constant c
    # Orientation provides Programming Tutorial 5) Slutzkys Theorem
    X_sigma <- sqrt(X_var) * abs(c)
    # Add both parameters to one dataframe for ease of showing
    x_c_df <- data.frame(matrix(NA, nrow = 1, ncol = 0))
    x_c_df$X_mu <- X_mu
    x_c_df$X_sigma <- X_sigma
    return(x_c_df)
  }
  
  # Now the loop trough the different as provided in the vector
  nf <- layout(matrix(c(1,2,3,4,5,6), ncol=2,nrow=3, TRUE),widths = c(1,1), heights=c(1.5,1.5,1.5))
  par(mar = c(2,4,1,0) + 0.1)
  for(a_ in a){
    # Draw the X_n and Y_n samples
    x_y_n <- get_x_y_n(a=a_, b=b, n=n, N_MC=N_MC)
    # Compute the part X*c for the a and b constellation
    # constant c has to be the set MEAN
    x_c <- get_x_c(a=a_, b=b, c=mean_y)
  
    # Compare by hist and density plot
    hist(x_y_n, probability=T, xlab="x-values",
         main=sprintf("Sim n:%s for N_MC:%s for a:%s and b:%s", n, N_MC, a_, b))
    # Add the density of the Hint, limiting normal distribution along the x values
    # with the specified parameters
    x_ <- seq(from=min(x_y_n), to=max(x_y_n), length.out=n)
    lines(x_, dnorm(x_, mean=x_c$X_mu, sd=x_c$X_sigma), col="red")
    abline(v=x_c$X_mu, lwd=3)
    legend("topright",
           legend = c("Lim Norm PDF", "mu X*c"),
           col = c("red", "black"),
           lty = c(1, 1),
           lwd=c(1,3))
  }
  par(mfrow = c(1, 1))
}

get_comp_plot(a, b, c, n, N_MC, mean_y, sd_y)
```